---
title: "Week Two Exercises"
author: "Chidi Justice"
date: "4/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE}
library(downloader)
library(tidyverse)

population <- read.csv("data/femaleControlsPopulation.csv")$Bodyweight
```

###### Data used in exercise has already been downloaded [here](week2Note.Rmd)  

##### Q1: What is the average of these weights?  
```{r}
mean(population) # 23.9 grams
```

##### Q2: After setting the seed at 1, set.seed(1) take a random sample of size 5. What is the absolute value (use abs) of the difference between the average of the sample and the average of all the values?  
```{r}
set.seed(1)
abs(mean(sample(population, 5)) - mean(population))
```

##### Q3: After setting the seed at 5, set.seed(5) take a random sample of size 5. What is the absolute value of the difference between the average of the sample and the average of all the values?  
```{r}
set.seed(5)
abs(mean(sample(population, 5)) - mean(population))
```